---
title: "Community Ecology Homework 1"
author: "Mary Van Dyke"
date: "2/28/2018"
output: pdf_document
---

1A 

```{r}
#set working directory
setwd("~/Box Sync/UCLA/Ecology200B")
#read in data
plot_data <- read.csv("plot_data.csv")
species_pool <- read.csv("species_pool.csv")
#subset species pool data to just include species from plot 25
temp <- subset(plot_data, plot_data$plot==25) # temporary data frame to get the species from plot 25
plot25 <- merge(temp, species_pool, by = "species") # merge with the species pool data from to get trait data for all the species in plot 25


```

Log Leaf Area Trait:
```{r}
# number of species in focal community/plot for null model:
N <- length(plot25$species)

## vector to store expected ranges of log leaf area from null communities:
lla_null_range<-NA

## number of iterations of null model to run
reps<- 999

logleafarea.vec <- c(species_pool$logLeafArea)

## loop to generate null communities and calculate expected trait ranges
for(ii in 1:reps){
  
  sample(logleafarea.vec, N)->lla_random_community
  
  lla_null_range[ii]<-abs(max(lla_random_community)-min(lla_random_community))
  
}

hist(lla_null_range, xlab= "range of log leaf area", ylab = "distribution from null model, N=6")

```


```{r}
hist(lla_null_range, xlab= "range of log leaf area", ylab = "distribution from null model, N=6", main= "Histogram of null model range of log leaf area", ylim = c(0, 270), xlim = c(0,4))

# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lla_null_range, probs=c(0.025, .975))->cutoffs
abline(v=cutoffs, lty=2)
text("2.5%", x=cutoffs[1], y=180, pos=4)
text("97.5%", x=cutoffs[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lla_null_range), lty=2)
text("mean", x=mean(lla_null_range), y=265, pos=4)

## add the observed range of log leaf area from plot 25
abs(max(plot25$logLeafArea) - min(plot25$logLeafArea))->lla_obs_range

abline(v=lla_obs_range, col='red', lwd=3)
text("observed value", x=lla_obs_range, y= 265, pos=4, col="red")

```

```{r}
## calculate a proper p value:

## how many null values is the observed value less than?
n_more <- sum(lla_obs_range<lla_null_range)
## how many null values is the observed value greater than?
n_less <-sum(lla_obs_range>lla_null_range)
## how many null values is the observed value tied with (this is especially an issue here because there are a small number of species in the pool- with larger pools ties are less common)
n_tied<-sum(lla_obs_range==lla_null_range)

## as we have a bunch of ties here, the convention is to rank your observed value in the middle of the ties. 
## p results (note that the tradition is also to add the observed value to the null distribution, hence 999 reps in the null and 999+1 in the denominator of the p value calculation)
print(c("one tailed test, range smaller than expected, p:", (n_less + n_tied/2) / (reps+1) ))
print(c("one tailed test, range greater than expected, p:", (n_more + n_tied/2) / (reps+1) ))
## two tailed tests are a little trickier:
if(n_less< ((reps+1)/2)){
  print(c("two tailed test, p:", (n_less + n_tied/2)*2 / (reps+1) ))
}else{
  print(c("two tailed test, p:", (n_more + n_tied/2)*2 / (reps+1) ))
}

```

```{r}
## calculate the standard effect size (s.e.s.)- a measure of the difference between the observed and expected value, expressed in terms of standard deviations of the null model:

#need mean and standard deviation of null distribution:
lla_null_mean<-mean(lla_null_range)
lla_null_sd<-sd(lla_null_range)

lla_range_SES<-(lla_obs_range - lla_null_mean)/lla_null_sd

print(c("standard effect size:", round(lla_range_SES, 3)))

```

Log Seed Size Trait

```{r}
## vector to store expected ranges of log leaf area from null communities:
lss_null_range<-NA

## number of iterations of null model to run
reps<- 999

## loop to generate null communities and calculate expected trait ranges
for(ii in 1:reps){
  
  sample(species_pool$logSeedSize, N)->lss_random_community
  
  lss_null_range[ii]<-abs(max(lss_random_community)-min(lss_random_community))
}

hist(lss_null_range, xlab= "range of log seed size", ylab = "distribution from null model, N=6", main= "Histogram of null model range of log seed size")


```


```{r}
hist(lss_null_range, xlab= "range of log seed size", ylab = "distribution from null model, N=6", main= "Histogram of null model range of log seed size")
# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lss_null_range, probs=c(0.025, .975))->cutoffs2
abline(v=cutoffs2, lty=2)
text("2.5%", x=cutoffs2[1], y=180, pos=4)
text("97.5%", x=cutoffs2[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lss_null_range), lty=2)
text("mean", x=mean(lss_null_range), y=170, pos=4)



## add the observed range of log seed size from plot 25
abs(max(plot25$logSeedSize) - min(plot25$logSeedSize))->lss_obs_range

abline(v=lss_obs_range, col='red', lwd=3)
text("observed value", x=lss_obs_range, y= 200, pos=4, col="red")

```


```{r}
## calculate a proper p value:

## how many null values is the observed value less than?
n_more2 <- sum(lss_obs_range<lss_null_range)
## how many null values is the observed value greater than?
n_less2 <-sum(lss_obs_range>lss_null_range)
## how many null values is the observed value tied with?
n_tied2<-sum(lss_obs_range==lss_null_range)

## two tailed tests:
if(n_less2< ((reps+1)/2)){
  print(c("two tailed test, p:", (n_less2 + n_tied2/2)*2 / (reps+1) ))
}else{
  print(c("two tailed test, p:", (n_more2 + n_tied2/2)*2 / (reps+1) ))
}

```

```{r}
## calculate the s.e.s.

#need mean and standard deviation of null distribution:
lss_null_mean<-mean(lss_null_range)
lss_null_sd<-sd(lss_null_range)

lss_range_SES<-(lss_obs_range - lss_null_mean)/lss_null_sd

print(c("standard effect size:", round(lss_range_SES, 3)))

```

Interpretation of Results:

For the leaf area trait it appears that the range in leaf areas in plot 25 is smaller than expected from the null model. The p-value for the two tailed t-test is less than 0.05 meaning the observed trait range was significantly different from the average null model range. It was more than 2 standard deviations away from the mean null range. The observed log seed size, however, was not significantly different from the mean null model log seed size range. This implies that there may be some environmental filtering that reduces the range of leaf areas in the particular area where plot 25 is. 

  
1B: Calculating null models for the same traits except this time every individual has an equal chance of dispersing a seed in the plot. 

Log Leaf Area Trait:  

```{r}
## vector to store expected ranges of log leaf area from null communities:
lla_null_range2<-NA

## loop to generate null communities and calculate expected trait ranges
for(ii in 1:reps){
  
  sample(species_pool$logLeafArea, N, prob = species_pool$abundance/sum(species_pool$abundance))->lla_random_community2
  
  lla_null_range2[ii]<-abs(max(lla_random_community2)-min(lla_random_community2))
  
}

hist(lla_null_range2, xlab= "range of log leaf area", ylab = "distribution from null model, N=6", main= "Histogram of null model range of log leaf area")
# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lla_null_range2, probs=c(0.025, .975))->cutoffs3
abline(v=cutoffs3, lty=2)
text("2.5%", x=cutoffs3[1], y=180, pos=4)
text("97.5%", x=cutoffs3[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lla_null_range2), lty=2)
text("mean", x=mean(lla_null_range2), y=265, pos=4)



## add the observed range of log leaf area from plot 25
abs(max(plot25$logLeafArea) - min(plot25$logLeafArea))->lla_obs_range

abline(v=lla_obs_range, col='red', lwd=3)
text("observed value", x=lla_obs_range, y= 265, pos=4, col="red")


```

```{r}
## calculate a proper p value:

## how many null values is the observed value less than?
n_more3 <- sum(lla_obs_range<lla_null_range2)
## how many null values is the observed value greater than?
n_less3 <-sum(lla_obs_range>lla_null_range2)
## how many null values is the observed value tied with (this is especially an issue here because there are a small number of species in the pool- with larger pools ties are less common)
n_tied3 <-sum(lla_obs_range==lla_null_range2)

## two tailed test:
if(n_less3 < ((reps+1)/2)){
  print(c("two tailed test, p:", (n_less3 + n_tied3/2)*2 / (reps+1) ))
}else{
  print(c("two tailed test, p:", (n_more3 + n_tied3/2)*2 / (reps+1) ))
}


```

```{r}
## calculate the standard effect size

#need mean and standard deviation of null distribution:
lla_null_mean2<-mean(lla_null_range2)
lla_null_sd2<-sd(lla_null_range2)

lla_range_SES2<-(lla_obs_range - lla_null_mean2)/lla_null_sd2

print(c("standard effect size:", round(lla_range_SES2, 3)))

```

Log Seed Size Trait:

```{r}
## vector to store expected ranges of log leaf area from null communities:
lss_null_range2<-NA

## loop to generate null communities and calculate expected trait ranges
for(ii in 1:reps){
  
  sample(species_pool$logSeedSize, N, prob = species_pool$abundance/sum(species_pool$abundance))->lss_random_community2
  
  lss_null_range2[ii]<-abs(max(lss_random_community2)-min(lss_random_community2))
  
}

hist(lss_null_range2, xlab= "range of log seed size", ylab = "distribution from null model, N=6", main= "Histogram of null model range of log seed size", xlim = c(0,7))

# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lss_null_range2, probs=c(0.025, .975))->cutoffs4
abline(v=cutoffs4, lty=2)
text("2.5%", x=cutoffs4[1], y=180, pos=4)
text("97.5%", x=cutoffs4[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lss_null_range2), lty=2)
text("mean", x=mean(lss_null_range2), y=170, pos=4)

abline(v=lss_obs_range, col='red', lwd=3) # line marking observed seed size range
text("observed value", x=lss_obs_range, y= 275, pos=4, col="red")

```

```{r}
## calculate a proper p value:

## how many null values is the observed value less than?
n_more4 <- sum(lss_obs_range<lss_null_range2)
## how many null values is the observed value greater than?
n_less4 <-sum(lss_obs_range>lss_null_range2)
## how many null values is the observed value tied with (this is especially an issue here because there are a small number of species in the pool- with larger pools ties are less common)
n_tied4<-sum(lss_obs_range==lss_null_range2)

## two tailed tests:
if(n_less4< ((reps+1)/2)){
  print(c("two tailed test, p:", (n_less4 + n_tied4/2)*2 / (reps+1) ))
}else{
  print(c("two tailed test, p:", (n_more4 + n_tied4/2)*2 / (reps+1) ))
}

```

```{r}
## calculate the s.e.s.

#need mean and standard deviation of null distribution:
lss_null_mean2<-mean(lss_null_range2)
lss_null_sd2<-sd(lss_null_range2)

lss_range_SES2<-(lss_obs_range - lss_null_mean2)/lss_null_sd2

print(c("standard effect size:", round(lss_range_SES2, 3)))

```


1C  
```{r}
##Subset the data to isolate plots by the river - 20, 21, and 28
temp20 <- subset(plot_data, plot_data$plot==20) # temporary data frame to get the species from plot 20
plot20 <- merge(temp20, species_pool, by = "species")
temp21 <- subset(plot_data, plot_data$plot==21) # temporary data frame to get the species from plot 21
plot21 <- merge(temp21, species_pool, by = "species")
temp28 <- subset(plot_data, plot_data$plot==28) # temporary data frame to get the species from plot 28
plot28 <- merge(temp28, species_pool, by = "species")
```

```{r}
#Create null model of log leaf areas to compare with plots 20, 21, 28. Since they all have a species N = 6 as does plot 25, you can use the same null model for them all. I am giving each individual an equal chance of dispersing into the random community so I am using the same null model I craeted for lof leaf area in part B. 

hist(lla_null_range2, xlab= "range of log leaf area", ylab = "distribution from null model, N=6", main= "Histogram of null model range of log leaf area for plot 20", xlim = c(0,4))
# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lla_null_range2, probs=c(0.025, .975))->cutoffs3
abline(v=cutoffs3, lty=2)
text("2.5%", x=cutoffs3[1], y=180, pos=4)
text("97.5%", x=cutoffs3[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lla_null_range2), lty=2)
text("mean", x=mean(lla_null_range2), y=265, pos=4)

## add the observed range of log leaf area from plot 20
max(plot20$logLeafArea) - min(plot20$logLeafArea) ->lla_obs_range_p20

abline(v=lla_obs_range_p20, col='red', lwd=3)
text("observed value", x=lla_obs_range_p20, y= 260, pos=4, col="red")


```
```{r}
## calculate a proper p value:

## how many null values is the observed value less than?
n_more_p20 <- sum(lla_obs_range_p20<lla_null_range2)
## how many null values is the observed value greater than?
n_less_p20 <-sum(lla_obs_range_p20>lla_null_range2)
## how many null values is the observed value tied with?
n_tied_p20 <-sum(lla_obs_range_p20==lla_null_range2)

## two tailed test:
if(n_less_p20 < ((reps+1)/2)){
  print(c("two tailed test, p value for plot 20 lla:", (n_less_p20 + n_tied_p20/2)*2 / (reps+1) ))
}else{
  print(c("two tailed test, p value for plot 20 lla:", (n_more_p20 + n_tied_p20/2)*2 / (reps+1) ))
}


```

```{r}
## calculate the standard effect size

#need mean and standard deviation of null distribution:
lla_null_mean2<-mean(lla_null_range2)
lla_null_sd2<-sd(lla_null_range2)

lla_range_SES_p20<-(lla_obs_range_p20 - lla_null_mean2)/lla_null_sd2

print(c("standard effect size:", round(lla_range_SES_p20, 3)))
```


```{r}
## Plot 21
hist(lla_null_range2, xlab= "range of log leaf area", ylab = "frequency", main= "Histogram of null model range of log leaf area, N=6", xlim = c(0,4))
# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lla_null_range2, probs=c(0.025, .975))->cutoffs3
abline(v=cutoffs3, lty=2)
text("2.5%", x=cutoffs3[1], y=180, pos=4)
text("97.5%", x=cutoffs3[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lla_null_range2), lty=2)
text("mean", x=mean(lla_null_range2), y=260, pos=4)

## add the observed range of log leaf area from plot 21
max(plot21$logLeafArea) - min(plot21$logLeafArea) ->lla_obs_range_p21

abline(v=lla_obs_range_p21, col='red', lwd=3)
text("plot 21 observed value", x=lla_obs_range_p21, y= 300, pos=4, col="red")

```


```{r}
## calculate a proper p value for plot 21 lla range:

## how many null values is the observed value less than?
n_more_p21 <- sum(lla_obs_range_p21<lla_null_range2)
## how many null values is the observed value greater than?
n_less_p21 <-sum(lla_obs_range_p21>lla_null_range2)
## how many null values is the observed value tied with?
n_tied_p21 <-sum(lla_obs_range_p21==lla_null_range2)

## two tailed test:
if(n_less_p21 < ((reps+1)/2)){
  print(c("two tailed test, p value for plot 21 lla:", (n_less_p21 + n_tied_p21/2)*2 / (reps+1) ))
}else{
  print(c("two tailed test, p value for plot 21 lla:", (n_more_p21 + n_tied_p21/2)*2 / (reps+1) ))
}

```

```{r}
##SES for plot 21 lla
lla_range_SES_p21<-(lla_obs_range_p21 - lla_null_mean2)/lla_null_sd2

print(c("standard effect size:", round(lla_range_SES_p21, 3)))

```

```{r}
#Create null model of log leaf areas for plot 28

hist(lla_null_range2, xlab= "range of log leaf area", ylab = "distribution from null model, N=7", main= "Histogram of null model range of log leaf area for plot 28", xlim = c(0,4))
# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lla_null_range2, probs=c(0.025, .975))->cutoffs3
abline(v=cutoffs3, lty=2)
text("2.5%", x=cutoffs3[1], y=180, pos=4)
text("97.5%", x=cutoffs3[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lla_null_range2), lty=2)
text("mean", x=mean(lla_null_range2), y=265, pos=4)

## add the observed range of log leaf area from plot 28
max(plot28$logLeafArea) - min(plot28$logLeafArea) ->lla_obs_range_p28

abline(v=lla_obs_range_p28, col='red', lwd=3)
text("observed value", x=lla_obs_range_p28, y= 300, pos=4, col="red")

```

```{r}
## calculate a proper p value for plot 28 lla range:

## how many null values is the observed value less than?
n_more_p28 <- sum(lla_obs_range_p28<lla_null_range2)
## how many null values is the observed value greater than?
n_less_p28 <-sum(lla_obs_range_p28>lla_null_range2)
## how many null values is the observed value tied with?
n_tied_p28 <-sum(lla_obs_range_p28==lla_null_range2)

## two tailed test:
if(n_less_p28 < ((reps+1)/2)){
  print(c("two tailed test, p value for plot 28 lla:", (n_less_p28 + n_tied_p28/2)*2 / (reps+1) ))
}else{
  print(c("two tailed test, p value for plot 28 lla:", (n_more_p28 + n_tied_p28/2)*2 / (reps+1) ))
}


```

```{r}
##SES for plot 28 lla
lla_range_SES_p28<-(lla_obs_range_p21 - lla_null_mean2)/lla_null_sd2

print(c("standard effect size:", round(lla_range_SES_p28, 3)))

```

Other way to answer this question:
```{r}
# Isolate all the species found in those three other plots near the stream - plot 25 doesn't include any other species that are not found in one of these three.
stream <-rbind(plot20, plot21, plot28)
stream$plot <- NULL # delete plot column
library(dplyr)
stream <- stream %>% distinct #delete repeated rows

```


```{r}
## Null model for community of stream plants
Nstream <- length(stream$species)
## vector to store expected ranges of log leaf area from null communities:
lla_stream_null<-NA

## loop to generate null communities and calculate expected trait ranges
for(ii in 1:reps){
  
  sample(species_pool$logLeafArea, Nstream, prob = species_pool$abundance/sum(species_pool$abundance))->lla_random_stream_comm
  
  lla_stream_null[ii]<-max(lla_random_stream_comm)-min(lla_random_stream_comm)
  
}

hist(lla_stream_null, xlab= "range of log leaf area", ylab = "frequency", main= "Histogram of null model range of log leaf area, N=16")

# quick *estimate* of sig. thresholds for plotting (for a two-tailed test)
quantile(lla_stream_null, probs=c(0.025, .975))->cutoffs5
abline(v=cutoffs5, lty=2)
text("2.5%", x=cutoffs5[1], y=180, pos=4)
text("97.5%", x=cutoffs5[2], y=180, pos=4)

## expected value (mean) from null:
abline(v=mean(lla_stream_null), lty=2)
text("mean", x=mean(lla_stream_null), y=170, pos=4)

#observed stream species range
max(stream$logLeafArea) - min(stream$logLeafArea) ->lla_obs_range_stream

abline(v=lla_obs_range_stream, col='red', lwd=3) # line marking observed lla range in stream plots
text("observed value", x=lla_obs_range_stream, y= 275, pos=4, col="red")

```

```{r}



```

```{r}


```

```{r}


```

```{r}


```


